---
title: "sketchpad_repeated"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggthemes)
library(lme4)
```

# Pilot 1: no feedback (had to wait until full drawing)

## Import group data

```{r}
d.pilot1.raw <- read_csv('../../../data/3dObjects/sketchpad_repeated/pilot1_group_data.csv')  %>% select(-png)
```

## Stroke histogram

```{r}
d.pilot1.raw %>% group_by(numStrokes) %>% ggplot(aes(x = numStrokes)) + geom_histogram()
```

## Find people who used crazy numbers of strokes
 
Also exclude if they clearly cheated (wrote word) 
```{r}
crazies.pilot1 <- unique((d.pilot1.raw %>% filter(condition == 'repeated' & numStrokes > 25) %>% select(gameID))$gameID)
crazies.pilot1 = c(crazies.pilot1, 
            '0766-fcb90e7e-bf4a-4a46-b6d6-3165b6c12b88', 
            '7024-8ac78089-539a-428b-9d0e-b52c71a0a1b4')
d.pilot1 <- d.pilot1.raw %>% filter(!(gameID %in% crazies.pilot1))

length(unique(d.pilot1$gameID))
```

```{r}
d.pilot1 %>%
  filter(condition == 'repeated') %>%
  filter()
  #filter(!(gameID %in% crazies)) %>%
  #filter(drawDuration < 60) %>%
  ggplot(aes(x=drawDuration)) + geom_histogram() + geom_vline(aes(xintercept = median(drawDuration)))
  # group_by(repetition, target, gameID) %>%
  # summarize(drawDuration = mean(drawDuration))
```

# Reduction by target 

```{r}
d.pilot1 %>%
  filter(condition == 'repeated') %>%
  group_by(repetition, target) %>%
  summarize(numStrokes = mean(drawDuration)) %>%
  ggplot(aes(x = repetition, y = numStrokes, group=target)) +
    geom_line() +
    theme_few() +
    geom_smooth(aes(group = 1), method ='loess')
```

# Reduction by gameid

```{r}
d.pilot1 %>%
  filter(condition == 'repeated') %>%
  group_by(repetition, gameID) %>%
  summarize(drawDuration = mean(drawDuration)) %>%
  ggplot(aes(x = repetition, y = drawDuration, group=interaction(gameID))) +
    geom_line(alpha=.3) +
    geom_smooth(aes(group = 1), method ='loess') +
    theme_few() +
    xlab('# repetitions') +
    ylab('# svg string length')
```

# Reduction (completely disaggregated)

```{r}
d.pilot1 %>%
  filter(condition == 'repeated') %>%
  #filter(!(gameID %in% crazies)) %>%
  #filter(drawDuration < 60) %>%
  group_by(repetition, target, gameID) %>%
  summarize(drawDuration = mean(drawDuration)) %>%
  ggplot(aes(x = repetition, y = drawDuration, group=interaction(gameID, target))) +
    geom_line(alpha=.1) +
    geom_smooth(aes(group = 1), method ='loess') +
    theme_few() +
    xlab('# repetitions') +
    ylab('draw duration') +
    ggtitle('Reduction per target per game') +
  #ylim(0, 25) +
  theme(aspect.ratio = 1)
```

## Mixed-effects model?

```{r}
summary(lmer(svgStringLength ~ repetition + (1 + repetition| gameID) + (1 + repetition| target), data = d %>%   
  filter(condition == 'repeated') %>%
  filter(!(gameID %in% crazies))))
```

## Compare repetition against control

```{r}
d.pilot1.prepost <- d.pilot1 %>%
  mutate(phase = case_when(
    trialType %in% c('0_control', '0_repeated') ~ 'pre',
    trialType %in% c('1_control', '7_repeated') ~ 'post'
  )) %>%
  mutate(phase = ordered(phase, levels = c('pre', 'post'))) %>%
  filter(!is.na(phase)) %>%
  # We don't know why but that 
  filter(gameID != "2904-d5fa6f4a-24e1-4568-b0ed-fb01ad503ed3") %>%
  group_by(phase, condition) %>%
  summarize(len = mean(drawDuration), se=sd(drawDuration)/sqrt(length(drawDuration)))

# d.toplot %>%
#   spread(condition, m) %>%
ggplot(d.pilot1.prepost, aes(x = phase, y = len, color = condition)) + #, linetype = condition, color=condition)) +
  geom_line(aes(group = condition)) + 
  geom_errorbar(aes(ymin=len-se,ymax=len+se), width=0) +
  #geom_text(aes(gameID))
  theme_few() +
  ylab('mean draw duration') +
  theme(aspect.ratio = 1) 
```

## prepost

```{r}
d.toplot <- d.pilot1 %>%
  mutate(phase = case_when(
    trialType %in% c('0_control', '0_repeated') ~ 'pre',
    trialType %in% c('1_control', '7_repeated') ~ 'post'
  )) %>%
  mutate(phase = ordered(phase, levels = c('pre', 'post'))) %>%
  filter(!is.na(phase)) %>%
  group_by(gameID, target, phase, condition) %>%
  summarize(len = mean(svgStringLength)) %>% #multi_boot_standard('svgStringLength')
  spread(phase, len) %>%
  mutate(reduction = post - pre) %>%
  group_by(gameID, condition) %>%
  summarize(m = mean(reduction))
  #multi_boot_standard('reduction')

d.toplot %>%
  spread(condition, m) %>%
ggplot(aes(x = control, y = repeated)) + #, linetype = condition, color=condition)) +
  #geom_bar(aes(group = condition)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  theme_few() +
  theme(aspect.ratio = 1)
```

# Pilot 2: real-time feedback (as soon as they knew)

```{r}
d.pilot2.raw <- read_csv('~/Repos/reference_games/data/3dObjects/sketchpad_repeated/feedback_pilot1_group_data.csv') %>%
  select(-png) 
```

### Stroke histogram

```{r}
d.pilot2.raw %>% ggplot(aes(x = numStrokes)) + geom_histogram(binwidth = 1) + theme_few()
```

### Find shift key users...

```{r}
crazies.pilot2 <- unique((d.pilot2.raw %>% filter(condition == 'repeated' & numStrokes > 25) %>% select(gameID))$gameID)
d.pilot2 <- d.pilot2.raw %>% filter(!(gameID %in% crazies.pilot2)) 
length(unique(d.pilot2$gameID))
```

### Num strokes over time

```{r}
d.pilot2 %>%
  filter(condition == 'repeated') %>%
  group_by(repetition, target) %>%
  summarize(numStrokes = mean(numStrokes)) %>%
  ggplot(aes(x = repetition, y = numStrokes, group=target)) +
    geom_line() +
    theme_few() +
    geom_smooth(aes(group = 1), method ='loess')
```

```{r}
d.pilot2 %>%
  filter(condition == 'repeated') %>%
  group_by(repetition, gameID) %>%
  summarize(numStrokes = mean(svgStringLength)) %>%
  ggplot(aes(x = repetition, y = numStrokes, group=gameID)) +
    geom_line() +
    theme_few() +
    geom_smooth(aes(group = 1), method ='loess')
```

```{r}
d.pilot2 %>%
  filter(condition == 'repeated') %>%
  group_by(repetition, target, gameID) %>%
  summarize(numStrokes = mean(svgStringLength)) %>%
  ggplot(aes(x = repetition, y = numStrokes, group=interaction(gameID, target))) +
    geom_line(alpha=.3) +
    geom_smooth(aes(group = 1), method ='loess') +
    theme_few() +
    xlab('# repetitions') +
    ylab('# strokes') +
    ggtitle('Reduction per target per game') +
  #ylim(0, 30) +
  theme(aspect.ratio = 1)
```

```{r}
d.pilot2 %>%
  filter(condition == 'repeated') %>%
  group_by(repetition, target, gameID) %>%
  summarize(drawDuration = mean(drawDuration)) %>%
  ggplot(aes(x = repetition, y = drawDuration, group=interaction(target, gameID))) +
    geom_line(alpha=.3) +
    geom_smooth(aes(group = 1), method ='loess') +
    theme_few() +
    xlab('# repetitions') +
    ylab('draw duration') +
    ggtitle('Reduction per target per game') +
  #ylim(0, 60) +
  theme(aspect.ratio = 1)
```

### Does accuracy go up?

```{r}
d.pilot2 %>%
  filter(condition == 'repeated') %>%
  group_by(trialNum) %>%
  mutate(outcome = ifelse(outcome == 'True', TRUE, FALSE)) %>%
  summarize(acc = mean(outcome)) %>%
  ggplot(aes(x = trialNum, y = acc)) +
    geom_point(alpha=.3) +
    geom_smooth(aes(group = 1), method ='loess') +
    theme_few() +
    xlab('# repetitions') +
    ylab('% acc') +
    ggtitle('Reduction per target per game') +
  ylim(0, 1) +
  theme(aspect.ratio = 1)
```

# Comparison of two pilots?

## Overall effect of feedback on accuracy?

```{r}
library(langcog)
rbind(d.pilot1 %>% mutate(src = 'draw-to-completition'),
      d.pilot2 %>% mutate(src = 'real-time feedback')) %>%
  filter(condition == 'repeated') %>%
  group_by(src, repetition) %>%
  mutate(outcome = ifelse(outcome == 'True', 1, 0)) %>%
  multi_boot_standard('outcome', nboot = 100) %>%
  ggplot(aes(x = repetition, y = mean, color = src)) +
    geom_line() +
    geom_errorbar(aes(ymax = summary_ci_upper, ymin = summary_ci_lower), width = 0) +
    theme_few() +
    ylab('accuracy') +
    xlab('repetition #') 
  
```

## Overall effect of feedback on draw duration?

```{r}
rbind(d.pilot1 %>% mutate(src = 'draw-to-completition'),
      d.pilot2 %>% mutate(src = 'real-time feedback')) %>%
  filter(condition == 'repeated') %>%
  filter(drawDuration < 60) %>%
  group_by(src, repetition) %>%
  summarize(m = mean(drawDuration), se = sd(drawDuration)/sqrt(length(drawDuration)))  %>%
  ggplot(aes(x = repetition, y = m, color = src)) +
    geom_line() +
    geom_errorbar(aes(ymax = m + se, ymin = m - se), width = 0) +
    theme_few() +
    ylab('draw duration (in seconds)') +
    xlab('repetition #')
```

## Overall effect of IES (summary statistic)

```{r}
rbind(d.pilot1 %>% mutate(src = 'draw-to-completition'),
      d.pilot2 %>% mutate(src = 'real-time feedback')) %>%
  filter(condition == 'repeated') %>%
  mutate(outcome = ifelse(outcome == 'True', 1, 0)) %>%
  group_by(src, gameID, repetition) %>%
  summarize(ies = mean(drawDuration)/mean(outcome))  %>%
  group_by(src, repetition) %>%
  multi_boot_standard('ies', nboot=1000)  %>%
  ggplot(aes(x = repetition, y = mean, color = src)) +
    geom_line() +
    geom_errorbar(aes(ymax = summary_ci_upper, ymin = summary_ci_lower), width = .1) +
    theme_few() +
    ylab('ies') +
    xlab('repetition #') +
    ylim(0, 60)
```

```{r}
numStrokes.bootstrapped = rbind(d.pilot1 %>% mutate(src = 'draw-to-completition'),
      d.pilot2 %>% mutate(src = 'real-time feedback')) %>%
  filter(condition == 'repeated') %>%
  mutate(outcome = ifelse(outcome == 'True', 1, 0)) %>%
  group_by(src, repetition) %>%
  #summarize(numStrokes = median(numStrokes))  %>%
  #group_by(src, repetition) %>%
  multi_boot_standard('numStrokes', nboot=1000) 

numStrokes.bootstrapped %>%
  ggplot(aes(x = repetition, y = mean, color = src)) +
    geom_line() +
    geom_errorbar(aes(ymax = summary_ci_upper, ymin = summary_ci_lower), width = .1) +
    theme_few() +
    ylab('mean # strokes') +
    xlab('repetition #') +
    ylim(0, 15)
```

```{r}
svgLength.bootstrapped <-rbind(d.pilot1 %>% mutate(src = 'draw-to-completition'),
      d.pilot2 %>% mutate(src = 'real-time feedback')) %>%
  filter(condition == 'repeated') %>%
  mutate(outcome = ifelse(outcome == 'True', 1, 0)) %>%
  group_by(src, repetition) %>%
  #summarize(numStrokes = median(numStrokes))  %>%
  #group_by(src, repetition) %>%
  multi_boot_standard('svgStringLength', nboot=1000) 

svgLength.bootstrapped %>%
  ggplot(aes(x = repetition, y = mean, color = src)) +
    geom_line() +
    geom_errorbar(aes(ymax = summary_ci_upper, ymin = summary_ci_lower), width = .1) +
    theme_few() +
    ylab('svgLength') +
    xlab('repetition #') 
    #ylim(0, 20)
```


## Pre-post

```{r}
d.pilot2.prepost <- d.pilot2 %>%
  mutate(phase = case_when(
    trialType %in% c('0_control', '0_repeated') ~ 'pre',
    trialType %in% c('1_control', '7_repeated') ~ 'post'
  )) %>%
  mutate(phase = ordered(phase, levels = c('pre', 'post'))) %>%
  filter(!is.na(phase)) %>%
  # We don't know why but that 
  group_by(phase, condition) %>%
  summarize(len = mean(drawDuration), se=sd(drawDuration)/sqrt(length(drawDuration)))

# d.toplot %>%
#   spread(condition, m) %>%
ggplot(d.pilot2.prepost, aes(x = phase, y = len, color = condition)) + #, linetype = condition, color=condition)) +
  geom_line(aes(group = condition)) + 
  geom_errorbar(aes(ymin=len-se,ymax=len+se), width=0) +
  #geom_text(aes(gameID))
  theme_few() +
  ylab('mean draw duration') +
  theme(aspect.ratio = 1) 
```