---
title: "chairs_chatbox"
output:
  pdf_document: default
  html_document: default
date: "7/18/2017"
---

#Import data

```{r, message=F, warning=F}
knitr::opts_chunk$set(echo = TRUE)

library(tidyr)
library(dplyr)
library(ggplot2)

setwd("~/Documents/cocolab/reference_games/analysis")

msgs_filenames <- list.files(path="../data/3dObjects/chairs_chatbox/chatMessage", pattern=".csv", full.names=TRUE)
msgs <- do.call(rbind, lapply(msgs_filenames, read.csv, sep='\t'))
#colnames(msgs) <- c("iterationName","gameid","time","trialNum","intendedName","role","text","reactionTime","eventType", "temp1", "temp2")
#msgs <- subset(msgs, select=-c(temp1,temp2))

clks_filenames <- list.files(path="../data/3dObjects/chairs_chatbox/clickedObj", pattern=".csv", full.names=TRUE)
clks <- do.call(rbind, lapply(clks_filenames, read.csv, sep='\t'))
#colnames(clks) <- c("iterationName","gameid","time","trialNum","object1Name","object1SketcherLoc","object1ViewerLoc","object2Name","object2SketcherLoc","object2ViewerLoc","object3Name","object3SketcherLoc","object3ViewerLoc","intendedName","clickedName","correct","condition","eventType", "temp1","temp2")
#clks <- subset(clks, select=-c(temp1,temp2))
```

# Pre-processing

We want to look only at complete (or nearly-complete) games. 

```{r}

rawAggregated <- clks %>%
  left_join(msgs, by =c("gameid","trialNum"))

#add a word count column
rawAggregated$numWords <- sapply(strsplit(as.character(rawAggregated$text), " "), length)

goodGamesList <- unique((rawAggregated %>%
  group_by(gameid) %>%
  filter(length(unique(trialNum)) >= 40))$gameid)

goodGames <- filter(rawAggregated, gameid %in% goodGamesList)

```

#Preliminary histograms/bar graphs

Dependent Variables
-Histogram of word distribution
-distribution of num words in different conditions (3 different facets within the same figure)

Summary bar plots
1.words sent by speaker during round
      a. Most important: words in first message
      b. words in all messages from round
2.total words sent by listener during round by condition
3. Trial duration across conditions

```{r}
#number of trials per game
ggplot(data=rawAggregated %>%
         group_by(gameid) %>%
         slice(which.max(trialNum)),
       aes(x = gameid, y = trialNum)) + 
  geom_bar(stat = "identity") +
  ggtitle("Number of trials per game")

#word distribution
ggplot(data=goodGames, aes(numWords)) + 
  geom_histogram(binwidth=1) +
  ggtitle("Word distribution")

#distribution of words in different conditions
ggplot(data=goodGames, aes(numWords)) + 
  geom_histogram(binwidth=1) +
  facet_wrap(~condition) +
  ggtitle("Word distribution by condition")

#number of trials where listener replied by condition



#number of incorrect clicks by condition
length(which(clks$correct=="false" & clks$condition=="close"))
length(which(clks$correct=="false" & clks$condition=="split"))
length(which(clks$correct=="false" & clks$condition=="far"))


# ggplot(data=clks %>%
#          group_by(gameid, condition),
#        aes(x = condition, y = length(which(clks$correct=="false")))) + 
#   geom_bar(stat = "identity") +
#   ggtitle("Number of incorrect clicks")
```

Do people use more words on harder trials?

```{r}
#total words sent by speaker during round
#total words 

#***average num of words in first message***
ggplot(data=goodGames %>%
         filter(role=="speaker") %>% 
         group_by(gameid, condition, trialNum) %>%
         slice(which.min(time.x)) %>%
         summarise(n = sum(numWords)/length(numWords)) %>%
         group_by(condition) %>%
         summarize(numWords = mean(n), se = sd(n)/sqrt(length(n))),
       aes(x = condition, y = numWords)) +
  geom_bar(stat="identity") + 
  ggtitle("Average words in first message only")

#Do people use more words on harder trials?
ggplot(goodGames %>%
         filter(role=="speaker") %>%
         group_by(gameid, condition, trialNum) %>%
         summarise(n = sum(numWords)/length(numWords)) %>%
         group_by(condition) %>%
         summarize(numWords = mean(n), se = sd(n)/sqrt(length(n))),
       aes(x = condition, y = numWords)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = numWords + se, ymin = numWords - se),
                  position="dodge", width=0.25) +
  ylab("Average # words per message per round")+
  theme_bw() +
  ggtitle("Average words by context")


#How about characters?
# ggplot(goodGames %>%
#          filter(role=="speaker") %>%
#          group_by(gameid, condition, trialNum) %>%
#          summarise(n = sum(numChars)/length(numChars)) %>%
#          group_by(condition) %>%
#          summarize(numChars = mean(n), se = sd(n)/sqrt(length(n))),
#        aes(x = condition, y = numChars)) +
#   geom_bar(stat = "identity", position = "dodge") +
#   geom_errorbar(aes(ymax = numChars + se, ymin = numChars - se),
#                   position="dodge", width=0.25) +
#   ylab("Average # characters per message per round")+
#   theme_bw() +
#   ggtitle("Average characters by context")


```

#Linguistic Analyses

First, import POS tags and merge
```{r}
tagged <- read.csv("posTagged.csv", sep = "\t")

POSAggregated <- clks %>%
  left_join(tagged, by =c("gameid","trialNum"))

#filter 
goodGamesList <- unique((POSAggregated %>%
  group_by(gameid) %>%
  filter(length(unique(trialNum)) >= 40))$gameid)

goodGames <- filter(POSAggregated, gameid %in% goodGamesList)

```

Look at distribution of parts of speech

```{r}
#make a POS count dataframe
posNames <- c("nouns","prepositions","verbs","determiners","pronouns","adjectives","adverbs","numbers","others")
posCounts <- colSums(goodGames[,posNames])

posSums <- data.frame(pos=posNames, counts=unname(unname(posCounts)))

ggplot(data=posSums,
       aes(x = posNames, y = posCounts)) + 
  geom_bar(stat ="identity") +
  scale_x_discrete(labels = abbreviate) + 
  ggtitle("Overall Part of Speech Counts")

#want: the next two graphs as grouped bar graph. any ideas?? might have to refactor data
#Avg number of superlatives by condition
ggplot(goodGames %>%
         group_by(gameid, condition, trialNum) %>%
         summarise(n = sum(numSuper)/length(numSuper)) %>%
         group_by(condition) %>%
         summarize(numSuper = mean(n), se = sd(n)/sqrt(length(n))),
       aes(x = condition, y = numSuper)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = numSuper + se, ymin = numSuper - se),
                  position="dodge", width=0.25) +
  ylab("Average # superlatives per message per round")+
  theme_bw() +
  ggtitle("Average number of superlatives by context")

#Avg number of comparitives by condition
ggplot(goodGames %>%
         group_by(gameid, condition, trialNum) %>%
         summarise(n = sum(numComp)/length(numComp)) %>%
         group_by(condition) %>%
         summarize(numComp = mean(n), se = sd(n)/sqrt(length(n))),
       aes(x = condition, y = numComp)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = numComp + se, ymin = numComp - se),
                  position="dodge", width=0.25) +
  ylab("Average # comparitives per message per round")+
  theme_bw() +
  ggtitle("Average number of comparitives by context")

#Avg number of negations by condition
ggplot(goodGames %>%
         group_by(gameid, condition, trialNum) %>%
         summarise(n = sum(negations)/length(negations)) %>%
         group_by(condition) %>%
         summarize(negations = mean(n), se = sd(n)/sqrt(length(n))),
       aes(x = condition, y = negations)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = negations + se, ymin = negations - se),
                  position="dodge", width=0.25) +
  ylab("Average # negations per message per round")+
  theme_bw() +
  ggtitle("Average number of negations by context")

```

Let's look at how language conventions change over the course of each game:

```{r}
#reduction analyses:
num_bins = 3
goodGames$trialBins <- cut(goodGames$trialNum, num_bins)

#number of words over time
ggplot(data=goodGames %>%
         group_by(gameid, trialBins) %>%
         summarise(n = sum(numWords)/length(numWords)) %>%
         group_by(trialBins) %>%
         summarize(numWords = mean(n), se = sd(n)/sqrt(length(n))),
       aes(x = trialBins, y = numWords)) +
  geom_bar(stat="identity") + 
  theme_bw() +
  ggtitle("Reduction: number of words over course of game")

#number of characters over time
ggplot(data=goodGames %>%
         group_by(gameid, trialBins) %>%
         summarise(n = sum(numChars)/length(numChars)) %>%
         group_by(trialBins) %>%
         summarize(numChars = mean(n), se = sd(n)/sqrt(length(n))),
       aes(x = trialBins, y = numChars)) +
  geom_bar(stat="identity") + 
  theme_bw() +
  ggtitle("Reduction: number of characters over course of game")

#the word "chair" over time
ggplot(data=goodGames %>%
         group_by(gameid, trialBins) %>%
         summarise(n = sum(w_chair)/length(w_chair)) %>%
         group_by(trialBins) %>%
         summarize(w_chair = mean(n), se = sd(n)/sqrt(length(n))),
       aes(x = trialBins, y = w_chair)) +
  geom_bar(stat="identity") + 
  theme_bw() +
  ggtitle("Reduction: the word 'chair' over course of game")

# #comparitives vs superlatives by condition
# ggplot(data=goodGames %>% 
#          group_by(trialBins),
#        aes(x = trialBins, y = numSuper)) + 
#   geom_bar(stat="identity") +
#   facet_wrap(~condition) + 
#   theme_bw() + 
#   ggtitle("Reduction: superlatives by condition")
# 
# ggplot(data=goodGames %>% 
#          group_by(trialBins),
#        aes(x = trialBins, y = numComp)) + 
#   geom_bar(stat="identity") +
#   facet_wrap(~condition) + 
#   theme_bw() + 
#   ggtitle("Reduction: Superlatives by condition")
  
#negations by condition
# ggplot(data=goodGames %>% 
#          group_by(trialBins),
#        aes(x = trialBins, y = numComp)) + 
#   geom_bar(stat="identity") +
#   facet_wrap(~condition) + 
#   theme_bw() + 
#   ggtitle("Negations by condition")


#adjectives vs prepositional phrases vs bare nouns by condition
#TODO

#line plot: all parts of speech 
posOfInterest <- c("nouns","prepositions","verbs","determiners","pronouns","adjectives","adverbs")

binNouns <- goodGames %>%
         group_by(gameid, trialBins) %>%
         summarise(n = sum(nouns)/length(nouns)) %>%
         group_by(trialBins) %>%
         summarize(counts = mean(n), se = sd(n)/sqrt(length(n)))
binPrepositions <- goodGames %>%
         group_by(gameid, trialBins) %>%
         summarise(n = sum(prepositions)/length(prepositions)) %>%
         group_by(trialBins) %>%
         summarize(counts = mean(n), se = sd(n)/sqrt(length(n)))
binVerbs <- goodGames %>%
         group_by(gameid, trialBins) %>%
         summarise(n = sum(verbs)/length(verbs)) %>%
         group_by(trialBins) %>%
         summarize(counts = mean(n), se = sd(n)/sqrt(length(n)))
binDeterminers <- goodGames %>%
         group_by(gameid, trialBins) %>%
         summarise(n = sum(determiners)/length(determiners)) %>%
         group_by(trialBins) %>%
         summarize(counts = mean(n), se = sd(n)/sqrt(length(n)))
binPronouns <- goodGames %>%
         group_by(gameid, trialBins) %>%
         summarise(n = sum(pronouns)/length(pronouns)) %>%
         group_by(trialBins) %>%
         summarize(counts = mean(n), se = sd(n)/sqrt(length(n)))
binAdjectives <- goodGames %>%
         group_by(gameid, trialBins) %>%
         summarise(n = sum(adjectives)/length(adjectives)) %>%
         group_by(trialBins) %>%
         summarize(counts = mean(n), se = sd(n)/sqrt(length(n)))
binAdverbs <- goodGames %>%
         group_by(gameid, trialBins) %>%
         summarise(n = sum(adverbs)/length(adverbs)) %>%
         group_by(trialBins) %>%
         summarize(counts = mean(n), se = sd(n)/sqrt(length(n)))

ggplot() +
  geom_point(data=binNouns, aes(x = trialBins, y = counts, group = 1)) + 
  geom_line(data=binNouns, aes(x = trialBins, y = counts, group = 1, color = "noun")) + 
  geom_point(data=binPrepositions, aes(x = trialBins, y = counts, group = 1)) + 
  geom_line(data=binPrepositions, aes(x = trialBins, y = counts, group = 1, color = "preposition")) + 
  geom_point(data=binVerbs, aes(x = trialBins, y = counts, group = 1)) + 
  geom_line(data=binVerbs, aes(x = trialBins, y = counts, group = 1, color = "verbs")) + 
  geom_point(data=binDeterminers, aes(x = trialBins, y = counts, group = 1)) + 
  geom_line(data=binDeterminers, aes(x = trialBins, y = counts, group = 1, color = "determiners")) + 
  geom_point(data=binAdjectives, aes(x = trialBins, y = counts, group = 1)) + 
  geom_line(data=binAdjectives, aes(x = trialBins, y = counts, group = 1, color = "adjectives")) + 
  geom_point(data=binAdverbs, aes(x = trialBins, y = counts, group = 1)) + 
  geom_line(data=binAdverbs, aes(x = trialBins, y = counts, group = 1, color = "adverbs")) + 
  scale_color_discrete(name = "part of speech") + 
  ggtitle("Reduction: all parts of speech")

  #normalize?
  
  
```

```{r}


```