---
title: "chairs_chatbox"
date: "7/18/2017"
output: html_document
---

#Import data

```{r, message=F, warning=F}
knitr::opts_chunk$set(echo = TRUE)

library(tidyr)
library(dplyr)
library(ggplot2)

setwd("~/Documents/cocolab/reference_games/analysis")

msgs_filenames <- list.files(path="../data/3dObjects/chairs_chatbox/chatMessage", pattern=".csv", full.names=TRUE)
msgs <- do.call(rbind, lapply(msgs_filenames, read.csv, sep='\t'))
#colnames(msgs) <- c("iterationName","gameid","time","trialNum","intendedName","role","text","reactionTime","eventType", "temp1", "temp2")
#msgs <- subset(msgs, select=-c(temp1,temp2))

clks_filenames <- list.files(path="../data/3dObjects/chairs_chatbox/clickedObj", pattern=".csv", full.names=TRUE)
clks <- do.call(rbind, lapply(clks_filenames, read.csv, sep='\t'))
#colnames(clks) <- c("iterationName","gameid","time","trialNum","object1Name","object1SketcherLoc","object1ViewerLoc","object2Name","object2SketcherLoc","object2ViewerLoc","object3Name","object3SketcherLoc","object3ViewerLoc","intendedName","clickedName","correct","condition","eventType", "temp1","temp2")
#clks <- subset(clks, select=-c(temp1,temp2))
```

# Pre-processing

We want to look only at complete (or nearly-complete) games. 

```{r}

rawAggregated <- clks %>%
  left_join(msgs, by =c("gameid","trialNum"))

#add a word count column
rawAggregated$numWords = sapply(strsplit(as.character(rawAggregated$text), " "), length)

goodGames <- unique((rawAggregated %>%
  group_by(gameid) %>%
  filter(length(unique(trialNum)) >= 40))$gameid)



```

#Preliminary histograms/bar graphs

Dependent Variables
-Histogram of word distribution
-distribution of num words in different conditions (3 different facets within the same figure)

Summary bar plots
1.words sent by speaker during round
      a. Most important: words in first message
      b. words in all messages from round
2.total words sent by listener during round by condition
3. Trial duration across conditions

```{r}
#number of trials per game
ggplot(data=rawAggregated %>%
         group_by(gameid) %>%
         slice(which.max(trialNum)),
       aes(x = gameid, y = trialNum)) + 
  geom_bar(stat = "identity") +
  ggtitle("Number of trials per game")

#word distribution
ggplot(data=rawAggregated, aes(numWords)) + 
  geom_histogram(binwidth=1) +
  ggtitle("Word distribution")

#distribution of words in different conditions
ggplot(data=rawAggregated, aes(numWords)) + 
  geom_histogram(binwidth=1) +
  facet_wrap(~condition) +
  ggtitle("Word distribution by condition")

#number of trials where listener replied by condition



#number of incorrect clicks by condition
length(which(clks$correct=="false" & clks$condition=="close"))
length(which(clks$correct=="false" & clks$condition=="split"))
length(which(clks$correct=="false" & clks$condition=="far"))


ggplot(data=clks %>%
         group_by(gameid, condition),
       aes(x = condition, y = length(which(clks$correct=="false")))) + 
  geom_bar(stat = "identity") +
  ggtitle("Number of incorrect clicks")
```


#Anaylsis (preliminary)

Do people use more words on harder trials?

```{r}

#total words sent by speaker during round
#total words 

#***average num of words in first message***
ggplot(data=rawAggregated %>%
         filter(role=="speaker") %>% 
         group_by(gameid, condition, trialNum) %>%
         slice(which.min(time.x)) %>%
         summarise(n = sum(numWords)/length(numWords)) %>%
         group_by(condition) %>%
         summarize(numWords = mean(n), se = sd(n)/sqrt(length(n))),
       aes(x = condition, y = numWords)) +
  geom_bar(stat="identity") + 
  ggtitle("Average words in first message only")

#Do people use more words on harder trials?
ggplot(rawAggregated %>%
         filter(role=="speaker") %>%
         group_by(gameid, condition, trialNum) %>%
         summarise(n = sum(numWords)/length(numWords)) %>%
         group_by(condition) %>%
         summarize(numWords = mean(n), se = sd(n)/sqrt(length(n))),
       aes(x = condition, y = numWords)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = numWords + se, ymin = numWords - se),
                  position="dodge", width=0.25) +
  ylab("Average # words per message per round")+
  theme_bw() +
  ggtitle("Average words by context")

```